---
title: "MUSA5000 Assignment 1"
author: "Anna Duan, Jingyi Li, and Jamie Song"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: yes
    toc: yes
    theme: cerulean
    toc_float: yes
    code_folding: hide
    number_sections: no
  pdf_document:
    toc: yes
---
# Introduction
```{r setup, include=TRUE, cache = TRUE, include = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

library(dplyr)
library(pander)
library(sf)
library(tidyr)
library(lubridate)
library(fuzzyjoin)
library(ggplot2)
library(tidygeocoder)
library(mapview)
library(tigris)
library(FNN)
library(janitor)
library(stringr)
library(ggpubr)
library(gridExtra)
library(ggcorrplot)
library(car)
library(MASS)
library(DAAG)

# Load themes for maps and plots (you can edit these)
mapTheme <- function(base_size = 12) {
  theme(
    text = element_text(color = "black", family="Helvetica"),
    plot.title = element_text(colour = "black", hjust=0),
    plot.subtitle=element_text(face="italic", hjust = 0),
    plot.caption=element_text(size = 7, hjust = 0),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.background = element_blank(),
    legend.position= c(0.8, 0.2)
  )
}

plotTheme <- function(base_size = 12) {
  theme(
    text = element_text(color = "black"),
    plot.title = element_text(colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size=12),
    axis.title = element_text(size=10),
    axis.text = element_text(size=8),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 10)
  )
} 
```
The ability to accurately predict house values is of great interest to real estate investors, homeowners and policymakers. This analysis examines the relationship between home sale values and select neighborhood attributes in Philadelphia.

In a survey of methods and input types for house price prediction, Geerts and colleagues (2023) present a set of variable types for prediction models, including structural features, temporal data, environmental features, and socioeconomic features. For this analysis, we focus on variables from the latter category, specifically the following:
- Residents with at least a bachelor’s degree (%)
- Housing units that are vacant (%)
- Housing units that are detached single family houses (%)
- Number of households with incomes below 100% poverty level
- Median household income (%).  
These indicators present a crude socioeconomic context for a neighborhood, and are well-documented as predictors of home prices. Indeed, it is reasonable that residents with higher educational attainment and income are able to afford more expensive homes. Conversely, it is also possible that the arrival of residents with higher socioeconomic status may lead to an eventual increase in property values. Using regression analysis, we find that __________________________. In light of this, we recommend that _________. Future research is needed _______________________.

# Methods  
## Data cleaning
The data in this analysis comes from the US Census Bureau's American Community Survey 5 year estimates. This dataframe is aggregated at the block-group level, with 1816 original observations. We clean the data to remove block groups with populations below 40, no housing units, median house values below $10,000, and one block-group with an extreme outlier for median house value. This leaves 1720 entries in our working dataframe. We additionally load a shapefile of Philadelphia's block-groups and join it to our dataframe for analysis and visualization. 

## Exploratory data analysis  
To prepare for our regression analysis, we first conduct an exploratory analysis by examining summary statistics as well as distributions of our independent and dependent variables. To test for co-linearity among our predictors, we also calculate the correlations between each independent variable. The correlation is a measure of how two variables change in relation to each other. The sample correlation coefficient, r, ranges from -1, indicating a perfect negative correlation, to 1, indicating a perfect positive correlation. A correlation coefficient of 0 indicates no relationship between two variables. The sample correlation coefficient r is calculated as follows:

[INSERT FORMULA HERE]

First, the mean values of both variables x and y are calculated. Following, we find both variables' standard deviations. For each of the n observations in the dataset, we subtract the mean of each variable (__, __) from the observation, and divide the difference (__, __) by the variable's standard deviation (__, __). We then multiply this quotient (__, __) for each observation's x and y variable and sum this product at each observation. Finally, we divide this sum by the number of observations minus 1. 

## Multiple regression analysis
For our regression analysis, we use a multiple 
```{r read data, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
dat <- read.csv("RegressionData.csv")

pander::pander(head(dat %>% dplyr::select(AREAKEY, MEDHVAL, PCTBACHMOR, MEDHHINC, PCTVACANT, PCTSINGLES,
                                    NBELPOV100)), caption = "Variables of interest")
  
dat.sf <- st_read("regression data shapefile/RegressionData.shp")
```

# Exploratory data analysis  
## Summary statistics
To begin our analysis, we first examine the summary statistics and distributions of our variables. The median home value (MEDHVAL) has a median value of $53,250 and a mean value of $66,288. This indicates that the data is positively skewed, with a higher share of observations below than the mean. Similarly, the independent variables (PCTBACHMOR, MEDHHINC, PCTVACANT, PCTSINGLES, and NBELPOV100) all have lower medians than means, indicating positive skewness.  
```{r summary statistics, message=FALSE, warning=FALSE, cache=TRUE, include=TRUE}
sum_stat <- function(var) {
  dat <- 
  dat %>%
    dplyr::select(var) %>%
    mutate(variable = var) %>%
    summarize(quartile_1 = quantile(dat[var], probs = 0.25, na.rm = TRUE),
           median = quantile(dat[var], probs = 0.5, na.rm = TRUE),
           mean = mean(dat[[var]], na.rm = TRUE),
           quartile_3 = quantile(dat[var], probs = 0.75, na.rm = TRUE),
           max = max(dat[var], na.rm = TRUE),
           sd = sd(dat[[var]], na.rm = TRUE),
           variance = var(dat[[var]], na.rm = TRUE))
  return(dat)
  }

medhval_summ <- sum_stat("MEDHVAL")
pctbachmorr_summ <- sum_stat("PCTBACHMOR")
medhhinc_summ <- sum_stat("MEDHHINC")
pctvacant_summ <- sum_stat("PCTVACANT")
pctsingles_summ <- sum_stat("PCTSINGLES")
nbelpov100_summ <- sum_stat("NBELPOV100")

dat.stat <- rbind(medhval_summ,
                  pctbachmorr_summ,
                  medhhinc_summ,
                  pctvacant_summ,
                  pctsingles_summ,
                  nbelpov100_summ) 

pander(dat.stat, caption = "Summary Statistics")
```

## Variable distributions 
Looking at the variables plotted as histograms, the positive skews are evident for median home value. individuals with bachelors degrees, vacant houses, single family houses, and households in poverty. Median household income is not visibly skewed. 
```{r histograms, fig.height=9, fig.width=13, message=FALSE, warning=FALSE, cache=TRUE}

grid.arrange(
ggplot(dat %>% filter(MEDHVAL <= 152525), aes(x = MEDHVAL)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "orchid", bins = 50) +
  labs(title = "Median Home Value ($)", y = "", x = "Dollars") +
  plotTheme(),

ggplot(dat %>% filter(PCTBACHMOR <= 59.13698), aes(x = PCTBACHMOR)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Individuals with Bachelor’s Degrees or Higher (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat %>% filter(MEDHHINC <= 56444.9), aes(x = MEDHHINC)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Median Household Income", y = "Block groups (n = 1720)", x = "Dollars") +
  plotTheme(),

ggplot(dat %>% filter(PCTVACANT <= 28.5154), aes(x = PCTVACANT)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Vacant Houses (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat %>% filter(PCTSINGLES <= 30.68176), aes(x = PCTSINGLES)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Single House Units (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat %>% filter(NBELPOV100 <= 514.05), aes(x = NBELPOV100)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Households Living in Poverty", y = "", x = "Households") +
  plotTheme(),
top = "Histogram of analysis variables")
```

## Log-transformed variable distributions 
After log-transformation, the dependent variable (Median Home Value) has a roughly normal distribution, so we will use LNMEDHVAL in our analysis. Of the independent variables, log-transformation only normalizes NBELPOV100 (Households living in poverty), so we will only use the log-transformed values for this variable and un-transformed values for the others.
```{r log transform, fig.height=9, fig.width=13, message=FALSE, warning=FALSE, cache=TRUE}
dat.log <- dat %>%
  mutate(LNMEDHVAL = log(MEDHVAL),
         LNPCTBACHMOR = log(1 + PCTBACHMOR), #adding one for variables with 0 values
         LNMEDHHINC = log(MEDHHINC),
         LNPCTVACANT = log(1 + PCTVACANT),
         LNPCTSINGLES = log(1 + PCTSINGLES),
         LNNBELPOV100 = log(1 + NBELPOV100))

grid.arrange(
ggplot(dat.log, aes(x = LNMEDHVAL)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "orchid", bins = 50) +
  labs(title = "Median Home Value ($)", y = "", x = "Dollars") +
  plotTheme(),

ggplot(dat.log, aes(x = LNPCTBACHMOR)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Individuals with Bachelor’s Degrees or Higher (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat.log, aes(x = LNMEDHHINC)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Median Household Income", y = "Block groups (n = 1720)", x = "Dollars") +
  plotTheme(),

ggplot(dat.log, aes(x = LNPCTVACANT)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Vacant Houses (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat.log, aes(x = LNPCTSINGLES)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Single House Units (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat.log, aes(x = LNNBELPOV100)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Households Living in Poverty", y = "", x = "Households") +
  plotTheme(),
top = "Histogram of log-transformed independent variables")
```

## Scatter plots
```{r scatter, fig.height=7, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
cor.long <- dat.log %>%
  dplyr::select(LNMEDHVAL, PCTBACHMOR, MEDHHINC, PCTVACANT, PCTSINGLES, LNNBELPOV100) %>%
  gather(Variable, Value, -LNMEDHVAL) %>%
  mutate(Value = as.numeric(Value))

cor.cor <-
  cor.long %>%
    group_by(Variable) %>%
    summarize(correlation = cor(Value, LNMEDHVAL, use = "complete.obs"))

ggplot(cor.long, aes(Value, LNMEDHVAL)) +
  geom_point(size = 0.1, color = "navy", alpha = 0.4) +
  geom_text(data = cor.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1, size = 4) +
  geom_smooth(method = "lm", se = FALSE, color = "orchid", linewidth = 0.5) +
  facet_wrap(~Variable, ncol = 3, scales = "free") +
  plotTheme() + 
  labs(title = "Median Home Value as a Function of Independent Variables", subtitle = "Philadelphia, PA Block Groups; n = 1720") 
```

## Correlation matrix  
Median household income is strongly correlated with log-transformed number of households living in poverty. We should only include one to avoid multicollinearity.  
```{r corr matrix, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}

dat.corplot <- dat.log %>%
  dplyr::select(-POLY_ID, -AREAKEY, -MEDHVAL, -LNPCTBACHMOR, -LNMEDHHINC, -LNPCTVACANT, -LNPCTSINGLES, -NBELPOV100)

ggcorrplot(round(cor(dat.corplot %>% na.omit()), 1),
  p.mat = cor_pmat(dat.corplot %>% na.omit()),
  colors = c("navy", "white", "orchid1"),
  type="lower",
  insig = "blank",
  show.legend = FALSE,
  lab = TRUE) +  
    labs(title = "Log-transformed median home value correlation with independent variables") 
```

## Chloropleth maps
```{r maps, fig.height=11, fig.width=13, message=FALSE, warning=FALSE, cache=TRUE}
grid.arrange(
ggplot(dat.sf) +
  geom_sf(aes(fill = LNMEDHVAL), color = "transparent") +
  scale_fill_viridis_c(option = "A", direction = 1) +
  labs(title = "Log-transformed median home value") +
  theme_void() +
  theme(legend.position = c(0.8, 0.2)),

ggplot(dat.sf) +
  geom_sf(aes(fill = PCTVACANT), color = "transparent") +
  scale_fill_viridis_c(option = "A", direction = 1) +
  labs(title = "Vacant homes (%)") +
  theme_void() +
  theme(legend.position = c(0.8, 0.2)),

ggplot(dat.sf) +
  geom_sf(aes(fill = PCTSINGLES), color = "transparent") +
  scale_fill_viridis_c(option = "A", direction = 1) +
  labs(title = "Single home units (%)") +
  theme_void() +
  theme(legend.position = c(0.8, 0.2)),

ggplot(dat.sf) +
  geom_sf(aes(fill = PCTBACHMOR), color = "transparent") +
  scale_fill_viridis_c(option = "A", direction = 1) +
  labs(title = "Individuals with bachelors or more (%)") +
  theme_void() +
  theme(legend.position = c(0.8, 0.2)),

ggplot(dat.sf) +
  geom_sf(aes(fill = LNNBELPOV), color = "transparent") +
  scale_fill_viridis_b(option = "A", direction = 1) +
  labs(title = "Log-transformed households in poverty") +
  theme_void() +
  theme(legend.position = c(0.8, 0.2)), ncol = 3)
```

# Multiple Regression Analysis
## Regression results 
```{r reg, message=FALSE, warning=FALSE, cache=TRUE}
# Assuming there’s no severe multicollinearity, use the lm command to run
# the regression where LNMEDHVAL is the dependent variable and
# PCTVACANT, PCTSINGLES, PCTBACHMOR, and LNNBELPOV100 are
# predictors.
# 
# In your report, be sure to present the summary of the fit as well as the
# ANOVA table containing the regression and error sum of squares (use the
# summary and anova commands). The only thing you should be looking at in
# the output from the anova command is the error sum of squares, and not
# any of the p-values.
# 
# c. Use the fitted, residuals and rstandard commands to save the predicted
# values, residuals and standardized residuals, respectively.

reg <- lm(LNMEDHVAL ~ PCTVACANT + PCTSINGLES + PCTBACHMOR + LNNBELPOV100, data=dat.log)
summary(reg)
anova(reg)
#vif(reg)
dat.log$pred_values <- fitted(reg)
dat.log$residual_values <- residuals(reg)
dat.log$standard_residuals <- rstandard(reg)

```

# Regression Assumption Checks 
## Scatterplot of standardized residuals
```{r scatter_standard_resid, message=FALSE, warning=FALSE, cache=TRUE}
# Create a scatter plot with Standardized Residuals on the y-axis and
# Predicted Values on the x-axis. You will be asked to present this scatter plot
# in your report, so take a screenshot of it if you plan to use MS Word.

ggplot(dat.log, aes(pred_values, standard_residuals)) +
  geom_point(size = 0.5, color = "navy", alpha = 0.4) +
  #geom_text(data = cor.cor, aes(label = paste("r =", round(correlation, 2))),
  #          x=-Inf, y=Inf, vjust = 1.5, hjust = -.1, size = 4) +
  geom_smooth(method = "lm", se = FALSE, color = "orchid", linewidth = 0.5) +
  plotTheme() + 
  labs(title = "Standardized residuals") 
```

## Additional Models
```{r stepwise and AIC, message=FALSE, warning=FALSE, cache=TRUE}
#Use the step and step$anova commands in the MASS library to run stepwise
# regression and determine the best model based on the Akaike Information
# Criterion.

step <- stepAIC(reg, direction="both")
step$anova
```


## K-folds cross validation
```{r k-folds cv, echo=TRUE, message=FALSE, warning=FALSE, cache=TRUE}
#Perform k-fold cross-validation (in which k = 5) using the CVlm command in the
# DAAG library and calculate the root mean square error (RMSE). Then re-run the
# regression model only using PCTVACANT and MEDHHINC as predictors, and again
# perform k-fold cross-validation in which k = 5. You will be asked to present the
# RMSE of both this model and the original model in your report.

cv <- CVlm(data=dat.log, reg, m=5)
mse <- attr(cv, "ms")
rmse <- sqrt(mse)
rmse #yields 0.3664 = better model

reg2 <- lm(LNMEDHVAL ~ PCTVACANT + MEDHHINC, data=dat.log)
cv2 <- CVlm(data=dat.log, reg2, m=5)
mse2 <- attr(cv2, "ms")
rmse2 <- sqrt(mse2)
rmse2 #yields 0.4427

```

## Histogram and chloropleth map of residuals
```{r residuals, message=FALSE, warning=FALSE, cache=TRUE}
#histogram and a choropleth map of standardized regression
#residuals that you saved using the rstandard command earlier
ggplot(dat.log, aes(standard_residuals)) +
  geom_histogram(binwidth = 0.5, fill = "navy") +
  plotTheme() + 
  labs(title = "Histogram of Standardized Residuals")

dat.sf<-merge(dat.log,dat.sf,by="POLY_ID")%>%
  st_as_sf()

ggplot(dat.sf) +
  geom_sf(aes(fill = standard_residuals), color = "transparent") +
  scale_fill_viridis_b(option = "A", direction = 1) +
  labs(title = "Chloropleth map of Standardized Residuals") +
  theme_void()
```

# Discussion and Limitations 

