---
title: "MUSA5000 Assignment 1"
author: "Anna Duan, Jingyi Li, and Jamie Song"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: yes
    toc: yes
    theme: cerulean
    toc_float: yes
    code_folding: hide
    number_sections: no
  pdf_document:
    toc: yes
---
# Introduction
```{r setup, include=TRUE, cache = TRUE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(dplyr)
library(pander)
library(sf)
library(tidyr)
library(lubridate)
library(fuzzyjoin)
library(ggplot2)
library(tidygeocoder)
library(mapview)
library(tigris)
library(FNN)
library(janitor)
library(stringr)
library(ggpubr)
library(gridExtra)
library(ggcorrplot)
library(car)
library(MASS)
library(DAAG)

# Load themes for maps and plots (you can edit these)
mapTheme <- function(base_size = 12) {
  theme(
    text = element_text(color = "black", family="Helvetica"),
    plot.title = element_text(colour = "black", hjust=0),
    plot.subtitle=element_text(face="italic", hjust = 0),
    plot.caption=element_text(size = 7, hjust = 0),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.background = element_blank(),
    legend.position= c(0.8, 0.2)
  )
}

plotTheme <- function(base_size = 12) {
  theme(
    text = element_text(color = "black"),
    plot.title = element_text(colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size=12),
    axis.title = element_text(size=10),
    axis.text = element_text(size=8),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 10)
  )
} 
```

# Methods 
```{r read data, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
dat <- read.csv("RegressionData.csv")

pander::pander(head(dat %>% dplyr::select(AREAKEY, MEDHVAL, PCTBACHMOR, MEDHHINC, PCTVACANT, PCTSINGLES,
                                    NBELPOV100)), caption = "Variables of interest")
  
dat.sf <- st_read("regression data shapefile/RegressionData.shp")
```

# Exploratory data analysis  
## Summary statistics
```{r summary statistics, warning = FALSE, message = FALSE, cache = TRUE}
sum_stat <- function(var) {
  dat <- 
  dat %>%
    dplyr::select(var) %>%
    mutate(variable = var) %>%
    group_by(variable) %>%
    summarize(quartile_1 = quantile(dat[var], probs = 0.25, na.rm = TRUE),
           median = quantile(dat[var], probs = 0.5, na.rm = TRUE),
           mean = mean(dat[[var]], na.rm = TRUE),
           quartile_3 = quantile(dat[var], probs = 0.75, na.rm = TRUE),
           max = max(dat[var], na.rm = TRUE),
           sd = sd(dat[[var]], na.rm = TRUE)) 
  return(dat)
  }

medhval_summ <- sum_stat("MEDHVAL")
pctbachmorr_summ <- sum_stat("PCTBACHMOR")
medhhinc_summ <- sum_stat("MEDHHINC")
pctvacant_summ <- sum_stat("PCTVACANT")
pctsingles_summ <- sum_stat("PCTSINGLES")
nbelpov100_summ <- sum_stat("NBELPOV100")

dat.stat <- rbind(medhval_summ,
                  pctbachmorr_summ,
                  medhhinc_summ,
                  pctvacant_summ,
                  pctsingles_summ,
                  nbelpov100_summ) 

pander(dat.stat, caption = "Summary Statistics")
```

## Histograms 
```{r histograms, fig.height=9, fig.width=13, message=FALSE, warning=FALSE, cache=TRUE}

grid.arrange(
ggplot(dat %>% filter(MEDHVAL <= 152525), aes(x = MEDHVAL)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "orchid", bins = 50) +
  labs(title = "Median Home Value ($)", y = "", x = "Dollars") +
  plotTheme(),

ggplot(dat %>% filter(PCTBACHMOR <= 59.13698), aes(x = PCTBACHMOR)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Individuals with Bachelor’s Degrees or Higher (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat %>% filter(MEDHHINC <= 56444.9), aes(x = MEDHHINC)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Median Household Income", y = "Block groups (n = 1720)", x = "Dollars") +
  plotTheme(),

ggplot(dat %>% filter(PCTVACANT <= 28.5154), aes(x = PCTVACANT)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Vacant Houses (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat %>% filter(PCTSINGLES <= 30.68176), aes(x = PCTSINGLES)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Single House Units (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat %>% filter(NBELPOV100 <= 514.05), aes(x = NBELPOV100)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Households Living in Poverty", y = "", x = "Households") +
  plotTheme(),
top = "Histogram of independent variables")
```

## Log transformed histograms 
After log-transformation, the dependent variable (Median Home Value) has a roughly normal distribution, so we will use LNMEDHVAL in our analysis. Of the independent variables, log-transformation only normalizes NBELPOV100 (Households living in poverty), so we will only use the log-transformed values for this variable and un-transformed values for the others.
```{r log transform, fig.height=9, fig.width=13, message=FALSE, warning=FALSE, cache=TRUE}
dat.log <- dat %>%
  mutate(LNMEDHVAL = log(MEDHVAL),
         LNPCTBACHMOR = log(1 + PCTBACHMOR), #adding one for variables with 0 values
         LNMEDHHINC = log(MEDHHINC),
         LNPCTVACANT = log(1 + PCTVACANT),
         LNPCTSINGLES = log(1 + PCTSINGLES),
         LNNBELPOV100 = log(1 + NBELPOV100))

grid.arrange(
ggplot(dat.log, aes(x = LNMEDHVAL)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "orchid", bins = 50) +
  labs(title = "Median Home Value ($)", y = "", x = "Dollars") +
  plotTheme(),

ggplot(dat.log, aes(x = LNPCTBACHMOR)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Individuals with Bachelor’s Degrees or Higher (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat.log, aes(x = LNMEDHHINC)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Median Household Income", y = "Block groups (n = 1720)", x = "Dollars") +
  plotTheme(),

ggplot(dat.log, aes(x = LNPCTVACANT)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Vacant Houses (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat.log, aes(x = LNPCTSINGLES)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Single House Units (%)", y = "", x = "Percent") +
  plotTheme(),

ggplot(dat.log, aes(x = LNNBELPOV100)) + # filtering for outliers, keeping 95%
  geom_histogram(fill = "navy", bins = 50) +
  labs(title = "Households Living in Poverty", y = "", x = "Households") +
  plotTheme(),
top = "Histogram of log-transformed independent variables")
```

## Scatter plots
```{r scatter, fig.height=7, fig.width=10, message=FALSE, warning=FALSE, cache=TRUE}
cor.long <- dat.log %>%
  dplyr::select(LNMEDHVAL, PCTBACHMOR, MEDHHINC, PCTVACANT, PCTSINGLES, LNNBELPOV100) %>%
  gather(Variable, Value, -LNMEDHVAL) %>%
  mutate(Value = as.numeric(Value))

cor.cor <-
  cor.long %>%
    group_by(Variable) %>%
    summarize(correlation = cor(Value, LNMEDHVAL, use = "complete.obs"))

ggplot(cor.long, aes(Value, LNMEDHVAL)) +
  geom_point(size = 0.1, color = "navy", alpha = 0.4) +
  geom_text(data = cor.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1, size = 4) +
  geom_smooth(method = "lm", se = FALSE, color = "orchid", linewidth = 0.5) +
  facet_wrap(~Variable, ncol = 3, scales = "free") +
  plotTheme() + 
  labs(title = "Median Home Value as a Function of Independent Variables", subtitle = "Philadelphia, PA Block Groups; n = 1720") 
```

## Correlation matrix  
Median household income is strongly correlated with log-transformed number of households living in poverty. We should only include one to avoid multicollinearity.  
```{r corr matrix, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}

dat.corplot <- dat.log %>%
  dplyr::select(-POLY_ID, -AREAKEY, -MEDHVAL, -LNPCTBACHMOR, -LNMEDHHINC, -LNPCTVACANT, -LNPCTSINGLES, -NBELPOV100)

ggcorrplot(round(cor(dat.corplot %>% na.omit()), 1),
  p.mat = cor_pmat(dat.corplot %>% na.omit()),
  colors = c("navy", "white", "orchid1"),
  type="lower",
  insig = "blank",
  show.legend = FALSE,
  lab = TRUE) +  
    labs(title = "Log-transformed median home value correlation with independent variables") 
```

## Chloropleth maps
```{r maps, fig.height=11, fig.width=13, message=FALSE, warning=FALSE, cache=TRUE}
grid.arrange(
ggplot(dat.sf) +
  geom_sf(aes(fill = LNMEDHVAL), color = "transparent") +
  scale_fill_viridis_c(option = "A", direction = 1) +
  labs(title = "Log-transformed median home value") +
  theme_void() +
  theme(legend.position = c(0.8, 0.2)),

ggplot(dat.sf) +
  geom_sf(aes(fill = PCTVACANT), color = "transparent") +
  scale_fill_viridis_c(option = "A", direction = 1) +
  labs(title = "Vacant homes (%)") +
  theme_void() +
  theme(legend.position = c(0.8, 0.2)),

ggplot(dat.sf) +
  geom_sf(aes(fill = PCTSINGLES), color = "transparent") +
  scale_fill_viridis_c(option = "A", direction = 1) +
  labs(title = "Single home units (%)") +
  theme_void() +
  theme(legend.position = c(0.8, 0.2)),

ggplot(dat.sf) +
  geom_sf(aes(fill = PCTBACHMOR), color = "transparent") +
  scale_fill_viridis_c(option = "A", direction = 1) +
  labs(title = "Individuals with bachelors or more (%)") +
  theme_void() +
  theme(legend.position = c(0.8, 0.2)),

ggplot(dat.sf) +
  geom_sf(aes(fill = LNNBELPOV), color = "transparent") +
  scale_fill_viridis_b(option = "A", direction = 1) +
  labs(title = "Log-transformed households in poverty") +
  theme_void() +
  theme(legend.position = c(0.8, 0.2)), ncol = 3)
```

# Multiple Regression Analysis
## Regression results 
```{r reg}
# Assuming there’s no severe multicollinearity, use the lm command to run
# the regression where LNMEDHVAL is the dependent variable and
# PCTVACANT, PCTSINGLES, PCTBACHMOR, and LNNBELPOV100 are
# predictors.
# 
# In your report, be sure to present the summary of the fit as well as the
# ANOVA table containing the regression and error sum of squares (use the
# summary and anova commands). The only thing you should be looking at in
# the output from the anova command is the error sum of squares, and not
# any of the p-values.
# 
# c. Use the fitted, residuals and rstandard commands to save the predicted
# values, residuals and standardized residuals, respectively.

reg <- lm(LNMEDHVAL ~ PCTVACANT + PCTSINGLES + PCTBACHMOR + LNNBELPOV100, data=dat.log)
summary(reg)
anova(reg)
#vif(reg)
dat.log$pred_values <- fitted(reg)
dat.log$residual_values <- residuals(reg)
dat.log$standard_residuals <- rstandard(reg)

```

# Regression Assumption Checks 
## Scatterplot of standardized residuals
```{r scatter_standard_resid}
# Create a scatter plot with Standardized Residuals on the y-axis and
# Predicted Values on the x-axis. You will be asked to present this scatter plot
# in your report, so take a screenshot of it if you plan to use MS Word.

ggplot(dat.log, aes(pred_values, standard_residuals)) +
  geom_point(size = 0.2, color = "navy", alpha = 0.4) +
  #geom_text(data = cor.cor, aes(label = paste("r =", round(correlation, 2))),
  #          x=-Inf, y=Inf, vjust = 1.5, hjust = -.1, size = 4) +
  geom_smooth(method = "lm", se = FALSE, color = "orchid", linewidth = 0.5) +
  plotTheme() + 
  labs(title = "Standardized residuals") 
```

## Additional Models
```{r stepwise and AIC}
#Use the step and step$anova commands in the MASS library to run stepwise
# regression and determine the best model based on the Akaike Information
# Criterion.

step <- stepAIC(reg, direction="both")
step$anova
```


## K-folds cross validation
```{r k-folds cv}
#Perform k-fold cross-validation (in which k = 5) using the CVlm command in the
# DAAG library and calculate the root mean square error (RMSE). Then re-run the
# regression model only using PCTVACANT and MEDHHINC as predictors, and again
# perform k-fold cross-validation in which k = 5. You will be asked to present the
# RMSE of both this model and the original model in your report.

cv <- CVlm(data=dat.log, reg, m=5)
mse <- attr(cv, "ms")
rmse <- sqrt(mse)
rmse #yields 0.3664 = better model

reg2 <- lm(LNMEDHVAL ~ PCTVACANT + MEDHHINC, data=dat.log)
cv2 <- CVlm(data=dat.log, reg2, m=5)
mse2 <- attr(cv2, "ms")
rmse2 <- sqrt(mse2)
rmse2 #yields 0.4427

```

## Histogram and chloropleth map of residuals
```{r residuals}
#histogram and a choropleth map of standardized regression
#residuals that you saved using the rstandard command earlier
```

# Discussion and Limitations 

